Reconstructing Propositional Proofs in Type Theory

### 2017-10-27 paper/sections/conclusions-and-future-work.tex

One of the main contribution of this study was
increasing the trustworthiness of the automatic prover \Metis.
Justifying a proof by a theorem prover
has a real significant impact for these automatic tools.
The reverse engineering task to understand the prover reasoning
can reveal important issues or bugs
in many parts of these systems (\eg, preprocessing, reasoning, or
deduction modules). During this research, we had the opportunity
to contribute to \Metis by reporting some bugs\footnote{Issues No.~2,
No.~4. and commit \name{8a3f11e} in the \Metis' official repository
\url{https://github.com/gilith/metis}}.
Fortunately, all these problems were fixed quickly by
\citeauthor{hurd2003first} in the \Metis 20170810 release.

>>>>

One of the main contribution of this study was increasing the
trustworthiness of the automatic prover \Metis.  Justifying a proof by
a theorem prover has a real significant impact for these automatic
tools.  The reverse engineering task to understand the prover
reasoning can reveal important issues or bugs in many parts of these
systems (\eg, preprocessing, reasoning, or deduction modules). During
this research, we had the opportunity to contribute to \Metis by
reporting some bugs---see Issues No.~2, No.~4, and commit
\name{8a3f11e} in \Metis' official
repository.\footnote{\url{https://github.com/gilith/metis}.}
Fortunately, all these problems were fixed quickly by Hurd in
\Metis~2.3 (release~20170822).

### 2017-10-27 paper/sections/metis-language-and-proofs.tex

\Metis is an automatic theorem prover for first-order logic with equality
developed by  \citeauthor{hurd2003first}~\cite{hurd2003first}. This prover is
suitable for proof-reconstruction since it provides well-documented proofs to
justify its deduction steps from the basis of only six inference rules for
first-order logic (see, for example, \cite{paulson2007source,Farber2015}). For
the propositional fragment, \Metis has three inference rules,
see Fig.~\ref{fig:metis-inferences}.

>>>>

\Metis is an automatic theorem prover for first-order logic with
equality developed by Hurd~\cite{hurd2003first}. This prover is
suitable for proof-reconstruction since it provides well-documented
proofs to justify its deduction steps from the basis of only six
inference rules for first-order logic (see, for example,
\cite{paulson2007source,Farber2015}). For the propositional fragment,
\Metis has three inference rules, see Fig.~\ref{fig:metis-inferences}.

### 2017-10-27 paper/sections/notes/discussion.tex
Cosmetic changes.

This lack of the \Metis' documentation represents one of the first barriers to build proof-reconstruction tools since some inference rules
like \canonicalize or \simplify summarize many essential steps to deduce the conclusion, making them hard to justify and emulate
(see a comment related on the presenting of \name{Sledgehammer}
in \citeauthor{paulson2007source}~\cite{paulson2007source} and \citeauthor{Farber2015}~\cite{Farber2015}).

@@

\textit{Normalization.} We observed \clausify rule described above is preceded by \canonicalize rule when \Metis introduces at some point of the derivation an axiom or a definition of the problem. In that case, both rules aim to perform a common task for
automatic theorem provers, Clausification, mainly described on paper
by \citeauthor{Sutcliffe1996}~\cite{Sutcliffe1996}. When the
aforementioned case does not occur we argue \canonicalize transforms
proposition into another one that mainly matches with its negative
normal form, its conjunctive normal form or its disjunctive normal
form. In addition, \canonicalize could also perform other tasks like
removing logic constants (⊥, ⊤), definitions, and tautologies, or even reorder both conjunctions and disjunctions of the proposition.

>>>>

This lack of the \Metis' documentation represents one of the first
barriers to build proof-reconstruction tools since some inference
rules like \canonicalize or \simplify summarize many essential steps
to deduce the conclusion, making them hard to justify and emulate (see
a comment related on the presenting of \name{Sledgehammer}
in~\cite{paulson2007source} and ~\cite{Farber2015}).

@@

\textit{Normalization.} We observed \clausify rule described above is
preceded by \canonicalize rule when \Metis introduces at some point of
the derivation an axiom or a definition of the problem. In that case,
both rules aim to perform a common task for automatic theorem provers,
Clausification, mainly described on~\cite{Sutcliffe1996}. When the
aforementioned case does not occur we argue \canonicalize transforms
proposition into another one that mainly matches with its negative
normal form, its conjunctive normal form or its disjunctive normal
form. In addition, \canonicalize could also perform other tasks like
removing logic constants (⊥, ⊤), definitions, and tautologies, or even
reorder both conjunctions and disjunctions of the proposition.

### 2017-10-27 paper/sections/proof-reconstruction.tex
Numbered all the examples.

The proof-reconstruction approach proposed consists of a series
of steps similar to the workflow presented by
\citeauthor{sultana2015}~\cite{sultana2015}.
This process is a translation from a source
system to a target system. In our case, the system of origin, the
automatic theorem prover, is \Metis; the target system is a proof
assistant, \Agda. We choose \Agda, but another proof assistant with the
same support of type theory and inductive types could be used.

@@

\begin{myexample}
\Metis has a inference rule named conjunct.
We refer to this rule writing \conjunct.

>>>>

The proof-reconstruction approach proposed consists of a series of
steps similar to the workflow presented~\cite{sultana2015}.  This
process is a translation from a source system to a target system. In
our case, the system of origin, the automatic theorem prover, is
\Metis; the target system is a proof assistant, \Agda. We choose
\Agda, but another proof assistant with the same support of type
theory and inductive types could be used.

@@

\begin{myexamplenum}
\Metis has a inference rule named conjunct.
We refer to this rule writing \conjunct.

### 2017-10-27 paper/sections/proof-reconstruction/resolve.tex

To construct a disjunction $ψ$ from a disjunction $φ$, we have used ideas
from the description in \citeauthor{bohme2010}~\cite{bohme2010} to prove
equality between nested disjunctions.
We define the $\fsbuild_{∨}$ function in~\eqref{eq:strong-build-or}
that uses every disjunct from the source formula, $φ$, to build up the
target disjunction $ψ$.

>>>>

To construct a disjunction~$ψ$ from a disjunction~$φ$, we have used
ideas from the description in~\cite{bohme2010} to prove equality
between nested disjunctions.  We define the $\fsbuild_{∨}$ function
in~\eqref{eq:strong-build-or} that uses every disjunct from the source
formula, $φ$, to build up the target disjunction $ψ$.
